---
title: "Identifying Optimal Time Slots for Marketing Use"
author: "Josh Schechter, Naiyani Paladugu, Lincoln, Orellana"
date: "2025-03-18"
output:
  html_document:
    code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Work on this branch. Do not push your code. We will copy and paste and put the code into one persons local and they will push and merge to develop/ This is our writeup. Everybody should write at least 1000 words for their section, and also include their plots and the code for that section.

```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(ggcorrplot)
```

```{r}
intervals_df <- read.csv("C:/Users/JDS221/GWU/DATS6101/projects/DATS6101_Project/data/fitbit/timestep_data/intervals.csv")
intervals_df <- na.omit(intervals_df)
intervals_df$time_only <- ifelse(nchar(intervals_df$interval_start) >= 19,
                       substr(intervals_df$interval_start, 12, 19),
                       "00:00:00")
```

## Introduction

Marketing is an effective tool for expanding knowledge about a given product or service. However, these advertisements must be aired at a time when people are available to view them to reach the largest possible audience. Our research sets out to determine the best times of day to air advertisements based on the availability of individuals. Our proposal is that potential customers are most available to be susceptible to consuming advertisements when they are awake but also resting. We collected data from Fitbit watch users which track features of 35 individuals throughout the course of a month, each with a day and time index. Our goal is to use these features at each time step to determine what time of day individuals are most susceptible to consuming marketing advertisements. We will do this by determining features of interest from the dataset, creating a scoring metric for each time interval, and determining if those time slots yield significant differences from one another. Based on the scoring function we derived and the statistical tests we ran on it; we determined that certain times of day are better for airing advertisements than others.

## Data Preprocessing

Our original dataset collected numerous features over a 1-month period, however not all features were found to be useful for our use-case. In fact, many features were not able to effectively join on our time intervals, so we decided to simply exclude them altogether. Other features were logged by minute and were easily merged on the derived time intervals. These features were as follows: average sleep, average steps taken, average intensity—scaled 1-3, with 3 being the highest—and average calories. Features that could have merged, but we decided to exclude due to redundancy were average METs (direct relationship with calories) and average heart rate.

We then aimed to condense our dataset into fewer datapoints, as time steps of 1-minute intervals were unnecessarily large. What we decided was to reduce the datapoints by changing the time intervals from 1-minute to 10-minutes, with feature values representing the mean for 1-minute interval averages. This condensed the data into more useful time slots for potential advertisers to analyze.

To complete our data frame we wanted to add some more columns that we thought could help to place data into bins for simplistic visualization at a later time. These columns were time of day, and military hour of the day. At this point, we are left with 4530 observations, each representing a 10-minute interval.

Finally, we cleaned our data of missing points. There were missing data points for individuals-often in chunks-at arbitrary timesteps, where we assumed that the individual was not wearing their Fitbit at that time. We elected to remove such observations. Interestingly, every individual was missing datapoints between the hours of 2:00 AM to 3:00 AM on March 13th, 2016. This was due to a time shift for daylight savings, meaning the times technically did not exist, so we removed those missing datapoints. Ultimately any datapoint with missing data was removed from the dataset.
Displayed below is a summary of the final data frame.

```{r}
head(intervals_df)
summary(intervals_df)
```
```{r}
# Aggregate by hour
hourly_summary_df <- intervals_df %>%
  group_by(hour_military) %>%
  summarise(
    avg_steps = mean(avg_steps, na.rm = TRUE),
    avg_sleep = mean(avg_sleep, na.rm = TRUE),
    avg_intensity = mean(avg_intensity, na.rm = TRUE),
    avg_calories = mean(avg_calories, na.rm = TRUE),
  )

minutes_summary_df <- intervals_df
```

**Hourly Summary**
```{r}
summary(hourly_summary_df)
summary(minutes_summary_df)
```

## Exploratory Data Analysis (EDA)

**Initial time series plots**
```{r}
# 1. Average Sleep
ggplot(hourly_summary_df, aes(x = hour_military, y = avg_sleep)) +
  geom_line(size = 1.2) +
  geom_point() +
  labs(title = "Average Sleep by Hour of Day", x = "Hour", y = "Average Sleep") +
  theme_minimal()

# 2. Average Steps
ggplot(hourly_summary_df, aes(x = hour_military, y = avg_steps)) +
  geom_line(size = 1.2) +
  geom_point() +
  labs(title = "Average Steps by Hour of Day", x = "Hour", y = "Average Steps") +
  theme_minimal()

# 3. Average Cals
ggplot(hourly_summary_df, aes(x = hour_military, y = avg_calories)) +
  geom_line(size = 1.2) +
  geom_point() +
  labs(title = "Average Calories by Hour of Day", x = "Hour", y = "Average Calories") +
  theme_minimal()

# 4. Average Intensity
ggplot(hourly_summary_df, aes(x = hour_military, y = avg_intensity)) +
  geom_line(size = 1.2) +
  geom_point() +
  labs(title = "Average Intensity by Hour of Day", x = "Hour", y = "Average Intensity") +
  theme_minimal()
```

Before constructing our scoring function, we sought to look at the features available and see how those change over time. We put together time-series plots with 24 bins, one for each hour of the day, each point representing the average activity per minute from that hour of the day. From this, four plots were derived.

The first plot being, the average sleep per minute, for each hour of the day. As expected, we see a lower amount of sleep per minute during daylight hours, whereas that number goes up for nighttime.

The second plot is the average number of steps taken per minute, for each hour of the day. Once again, we see an expected result, where daylight hours end up being the most steps per minute, with a significant decline during nighttime hours.  This trend continues for the next two plots as well. The first of which being average calories burned per minute, for each hour of the day. The second being the average intensity per minute, for each hour of the day. Both plots were consistent with the expectation of seeing higher values during daylight hours, and lower values at night. By looking at our visualizations, we inferred that some of our features are very strongly correlated, and decided to investigate this observation.

**Correlation Matrix**
```{r}
correlation_matrix <- cor(intervals_df[, c("avg_steps", "avg_sleep", "avg_calories","avg_intensity")])
ggcorrplot(correlation_matrix, hc.order = TRUE, type = "lower", lab = TRUE)
```

After visualizing how individual features behave throughout the day, we determined that there could be some strong correlations between average steps, calories, and intensity. If this was the case, it would make sense to remove some of the redundant features, allowing us to reduce the dimensionality of our scoring function. 

After running the correlation matrix, our suspicions were confirmed. Average steps, calories, and intensity were all very strongly correlated to one another. Notably, sleep had a very weak correlation to any of the other features. To reduce the dimensionality, we decided to exclude steps and calories from our scoring function, as we felt that one of these three features would be sufficient at representing activity level. At this point, we have identified our features of interest as average sleep and intensity.


```{r}
minutes_summary_df <- minutes_summary_df %>%
  mutate(
    norm_sleep = (avg_sleep - min(avg_sleep)) / (max(avg_sleep) - min(avg_sleep)),
    norm_intensity = (avg_intensity - min(avg_intensity)) / (max(avg_intensity) - min(avg_intensity))
  )

hourly_summary_df <- hourly_summary_df %>%
  mutate(
    norm_sleep = (avg_sleep - min(avg_sleep)) / (max(avg_sleep) - min(avg_sleep)),
    norm_intensity = (avg_intensity - min(avg_intensity)) / (max(avg_intensity) - min(avg_intensity))
  )
```

**Break down opportunity Score**
```{r}
weight_intensity <- 0.4
weight_sleep <- 0.6

minutes_summary_df <- minutes_summary_df %>%
  mutate(
    opportunity_score = weight_intensity*(1 - abs(norm_intensity - 0.5) * 2) + weight_sleep*(1 - norm_sleep)
  )

hourly_summary_df <- hourly_summary_df %>%
  mutate(
    opportunity_score = weight_intensity*(1 - abs(norm_intensity - 0.5) * 2) + weight_sleep*(1 - norm_sleep)
  )
```

By identifying any lingering redundancies in our dataset, we were now able to create our scoring function with the features sleep and intensity.

Before creating our function, it was necessary to normalize—or scale—the features to support our scoring function. We found normalizing the data yielded better scores. After normalization our two new variables, norm_intensity, and norm_sleep represent the normalized datapoints for our intensity and sleep features.

Next, we created our scoring function, called opportunity score. The opportunity score formula is as follows:

opportunity_score = 0.4\*(1 - |norm_intensity – 0.5| \* 2) + 0.6\*(1 - norm_sleep)

Inside the first set of parentheses, is a formula to reward values closer to the middle of the distribution. Originally, we created a formula that rewards low activity, however that would heavily favor nighttime hours, when individuals are likely to be sleeping. To combat this, we used this formula to ensure that moderate intensity is favored. If there is moderate intensity, we can be more confident that an individual is not sleeping during that time interval but rather is resting.

Inside the second set of parentheses, is the normalized value of sleep per datapoint, subtracted from the normalized maximum value of one. This allows the score to best reward low sleep values.

Each set of parameters are multiplied by a weight value. The weights were determined by how important we felt each feature was to the score based on our use-case. We decided to weigh sleep more heavily than intensity, as an individual who is sleeping has no potential to consume a marketing advertisement, whereas a woke but active individual could still be exposed to marketing.


**Opportunity Score Results**
```{r}
summary(minutes_summary_df$opportunity_score)
mean(minutes_summary_df$opportunity_score)
sd(minutes_summary_df$opportunity_score)
var(minutes_summary_df$opportunity_score)
```
```{r}
summary(hourly_summary_df$opportunity_score)
mean(hourly_summary_df$opportunity_score)
sd(hourly_summary_df$opportunity_score)
var(hourly_summary_df$opportunity_score)
```
Displayed above are the summary and descriptive statistics of the newly derived feature opportunity score. The mean is a score of 0.230, the standard deviation is 0.175, and the variance is 0.031. When looking at the summary statistics, the first and third quartiles appear to be relatively low. These early indications suggest that the distribution will be bottom heavy, where most scores are relatively low. Moreover, there is potential for a right skew. As we go on, this is something that we felt was necessary to investigate and account for.

```{r}
hourly_summary_df <- minutes_summary_df %>%
  group_by(hour_military) %>%
  summarize(opportunity_score = mean(opportunity_score, na.rm = TRUE),
            norm_intensity = mean(norm_intensity, na.rm = TRUE),
            norm_sleep = mean(norm_sleep, na.rm = TRUE)
            )
            
```

```{r}
minutes_summary_df <- minutes_summary_df %>%
  mutate(
    opportunity_level = cut(opportunity_score,
                            breaks = quantile(opportunity_score, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE),
                            labels = c("bad", "mid", "good"),
                            include.lowest = TRUE)
  )

hourly_summary_df <- hourly_summary_df %>%
  mutate(
    opportunity_level = cut(opportunity_score,
                            breaks = quantile(opportunity_score, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE),
                            labels = c("bad", "mid", "good"),
                            include.lowest = TRUE)
  )
```

#More EDA (Lincoln)
 How did your question change, if at all, after Exploratory Data Analysis?   Based on EDA can you begin to sketch out an answer to your question?  Summary of the dataset  Descriptive statistics  Graphical representations of the data  [Proper usage when applicable] Measures of Variance / sd 

Understanding user behavior via sleep, intensity, and opportunity score metrics is crucial for optimizing advertising and engagement strategies. This EDA uses various boxplots, histograms, and statistical summaries to uncover trends in daily activity patters to help determine when users are most and least receptive to engagement.


## Boxplot Opportunity Score (Include descriptive statistics)
```{r}
# Average
ggplot(minutes_summary_df, aes(y = opportunity_score)) +
  geom_boxplot(fill = "lightblue", alpha = 0.7, outlier.color = "red") +
  labs(title = "Overall Boxplot of Opportunity Score", y = "Opportunity Score") +
  theme_minimal()

summary(intervals_df$opportunity_score)

# Hourly

ggplot(hourly_summary_df, aes(x = as.factor(hour_military), y = opportunity_score)) +
  geom_boxplot(fill = "lightblue", alpha = 0.7, outlier.color = "red") +
  labs(title = "Boxplot of Opportunity Score by Hour", x = "Hour (Military Time)", y = "Opportunity Score") +
  theme_minimal()

tapply(hourly_summary_df$opportunity_score, hourly_summary_df$hour_military, summary)
```

## Boxplot Intensity (Include descriptive statistics)
```{r}
# Average
ggplot(minutes_summary_df, aes(y = norm_intensity)) +
  geom_boxplot(fill = "lightcoral", alpha = 0.7, outlier.color = "red") +
  labs(title = "Overall Boxplot of Intensity", y = "Intensity") +
  theme_minimal()

summary(minutes_summary_df$avg_intensity)

# Hourly
ggplot(hourly_summary_df, aes(x = as.factor(hour_military), y = norm_intensity)) +
  geom_boxplot(fill = "lightcoral", alpha = 0.7, outlier.color = "red") +
  labs(title = "Boxplot of Intensity by Hour", x = "Hour (Military Time)", y = "Intensity") +
  theme_minimal()

tapply(hourly_summary_df$norm_intensity, hourly_summary_df$hour_military, summary)
```

Most users show low-to-moderate activity levels, suggesting sedentary behavior. When looking at the hourly boxplot, we see minimal activity being recorded between 00:00 and 05:00, aligning with sleep hours untila sudden increase at 06:00, suggesting morning workouts and routines. From 16:00-19:00, we observe yet another increase which can suggest evening comutes/routines. After 20:00, the intensity gradually declines to reinforce that late evening are more sedentary and show a higher chance of advertisements reaching those who aren't physically engaged.

## Boxplot Sleep (Include descriptive statistics)
```{r}
# Average
ggplot(minutes_summary_df, aes(y = norm_sleep)) +
  geom_boxplot(fill = "gold", alpha = 0.7, outlier.color = "red") +
  labs(title = "Overall Boxplot of Sleep", y = "Sleep") +
  theme_minimal()

summary(minutes_summary_df$avg_sleep)

# Hourly
ggplot(hourly_summary_df, aes(x = as.factor(hour_military), y = norm_sleep)) +
  geom_boxplot(fill = "gold", alpha = 0.7, outlier.color = "red") +
  labs(title = "Boxplot of Sleep by Hour", x = "Hour (Military Time)", y = "Sleep") +
  theme_minimal()

tapply(hourly_summary_df$norm_sleep, hourly_summary_df$hour_military, summary)
```
The overall average sleep boxplot is closest to 1, meaning users receive adequate sleep during expected hours. Some outliers are shown below 0.5, suggesting some users can either be disrupted during sleep or have irregular patterns. The hourly boxplot shows sleep is highest between 22:00 and 06:00, visualizing the transition from sleep to wakefulness. Logically, most people during the day would be doing some form of activity whether it'd be work, school, or exercise (hence why there are outliers between 10:00 and 16:00). Since the early morning between 06:00 and 08:00 shows a transitional period, it could be an optimal time for subtle engagement strategies, such as morning news briefings, podcasts, or passive advertisements.

## Bar Graphs
```{r}
# Mean opportunity score by time interval
ggplot(hourly_summary_df, aes(x = hour_military, y = opportunity_score)) +
  geom_col(fill = "darkorange") +
  labs(title = "Average Opportunity Score by 1 Minute Interval (Best Time for TV Ads)",
       x = "10 Minute Intervals (Military Time)",
       y = "Mean Opportunity Score") +
  theme(axis.text.x = element_blank())

intensity_summary <- minutes_summary_df %>%
  group_by(time_only) %>%
  summarize(norm_intensity = mean(norm_intensity, na.rm = TRUE))

# Mean intensity by time interval
ggplot(intensity_summary, aes(x = time_only, y = norm_intensity)) +
  geom_col(fill = "green") +
  labs(title = "Average Intensity by 1 Minute Interval (Best Time for TV Ads)",
       x = "10 Minute Intervals (Military Time)",
       y = "Mean Intensity") +
  theme(axis.text.x = element_blank())

sleep_summary <- minutes_summary_df %>%
  group_by(time_only) %>%
  summarize(norm_sleep = mean(norm_sleep, na.rm = TRUE))

# Mean sleep by time interval
ggplot(sleep_summary, aes(x = time_only, y = norm_sleep)) +
  geom_col(fill = "blue") +
  labs(title = "Average Sleep by 1 Minute Interval (Best Time for TV Ads)",
       x = "10 Minute Intervals (Military Time)",
       y = "Mean Sleep") +
  theme(axis.text.x = element_blank())
```

The bar graphs depicting average opportunity score, intensity, and sleep by 10-minute intervals provide a more granular view of activity trends throughout the day. The opportunity score peaks between mid-morning and late evening (8 AM – 10 AM and 8 PM – 11 PM), aligning with times when users are likely in a more passive state. The intensity bar graph further supports earlier findings—activity increases in the morning, peaks in the late afternoon, and declines into the evening. The sleep graph reinforces the expected cycle, with the lowest sleep scores occurring in daylight hours and reaching near-max levels at night.


## QQ Plot Opportunity Score (Make sure to label)
```{r}
qqnorm(minutes_summary_df$opportunity_score)
qqline(minutes_summary_df$opportunity_score, col = "red")

qqnorm(hourly_summary_df$opportunity_score)
qqline(hourly_summary_df$opportunity_score, col = "red")
```

## QQ Plot Intensity
```{r}
qqnorm(minutes_summary_df$norm_intensity)
qqline(minutes_summary_df$norm_intensity, col = "red")

qqnorm(hourly_summary_df$norm_intensity)
qqline(hourly_summary_df$norm_intensity, col = "red")
```


## QQ Plot Sleep
```{r}
qqnorm(minutes_summary_df$norm_sleep)
qqline(minutes_summary_df$norm_sleep, col = "red")

qqnorm(hourly_summary_df$norm_sleep)
qqline(hourly_summary_df$norm_sleep, col = "red")
```

The Q-Q plots for sleep, intensity, and opportunity score confirm that none of these variables are normally distributed. Instead, they exhibit significant skewness, meaning that user behavior is highly variable. This has major implications for predictive modeling, suggesting that a segmentation-based approach is needed rather than assuming a universal pattern of engagement.


## Histogram Distribution Plot Opportunity Score
```{r}
ggplot(minutes_summary_df, aes(x = opportunity_score)) +
  geom_histogram(binwidth = 0.1, fill = "lightgreen", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Opportunity Score", x = "Opportunity Score", y = "Frequency") +
  theme_minimal()
```

The histogram shows most scores falling between 0.3 and 0.5, with some peaks between 0.4 and 0.5 in the evening. The visualization of opportunity score as well suggests the best times for engagement being the morning transition (06:00-09:00), lunchtime (12:00-14:00), and evening (21:00-23:00)


## Histogram Intensity
```{r}
ggplot(minutes_summary_df, aes(x = norm_intensity)) +
  geom_histogram(binwidth = 0.05, fill = "purple", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Intensity", x = "Intensity", y = "Frequency") +
  theme_minimal()
```

## Histogram Sleep
```{r}
ggplot(minutes_summary_df, aes(x = norm_sleep)) +
  geom_histogram(binwidth = 0.05, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Sleep", x = "Average Sleep", y = "Frequency") +
  theme_minimal()
```

The histograms of intensity and sleep provide further context to our exploratory analysis. The intensity histogram reveals a right-skewed distribution, where most recorded values are low, confirming that users spend a significant portion of their time in low-movement states. The sleep histogram, in contrast, shows a strong left-skew, where most sleep values are clustered near 1.0, suggesting that the majority of users receive high sleep scores, but a small subset experiences irregular or fragmented sleep.



##LINCOLN -> MAKE SURE TO INCLUDE Measures of Variance / sd  IF NECESSARY
```{r}
summary_stats <- intervals_df %>%
  summarise(
    sd_sleep = sd(avg_sleep, na.rm = TRUE),
    iqr_sleep = IQR(avg_sleep, na.rm = TRUE),
    sd_steps = sd(avg_steps, na.rm = TRUE),
    iqr_steps = IQR(avg_steps, na.rm = TRUE),
    sd_intensity = sd(avg_intensity, na.rm = TRUE),
    iqr_intensity = IQR(avg_intensity, na.rm = TRUE),
    sd_calories = sd(avg_calories, na.rm = TRUE),
    iqr_calories = IQR(avg_calories, na.rm = TRUE)
  )
print(summary_stats)
```

The variance analysis provides crucial insights into the consistency and variability of user behaviors, reinforcing patterns observed in the exploratory data analysis (EDA). The low standard deviation and interquartile range (IQR) for sleep suggest that most users follow predictable sleep schedules, with minimal variation between individuals. This consistency aligns with previous findings that nighttime hours (10 PM – 6 AM) are dominated by sleep, making them less ideal for engagement strategies. Conversely, step counts and caloric expenditure display high variability, indicating that while some users are highly active, others remain largely sedentary. This confirms the need for targeted engagement strategies based on behavioral segmentation, as a one-size-fits-all approach would be ineffective.

Additionally, physical intensity levels exhibit low deviation, supporting earlier conclusions that most users engage in low to moderate activity throughout the day. This suggests that advertising and content strategies should be optimized for moments of low intensity, particularly during lunch breaks and late evening hours when users are more likely to engage passively. Ultimately, this variance analysis strengthens the overall project goal—identifying the best times for digital engagement based on behavioral trends, ensuring content is delivered when users are most receptive.

##LINCOLN -> MAKE SURE TO INCLUDE Measures of Variance / sd  IF NECESSARY

## Statistical Analysis (Naiyani)
**TODO:**
1. Run Kruskal Wallis on: Opportunity Score, Intensity, Sleep for both hourly and minutes
2. Only move on wit
2. Check for conditions to run chi square
3. Run Chi Square on Opportunity Score by 10-minute intervals only since it is proven significant
4. Run post hoc on Opportunity Score by 10-minute intervals only
5. Write up your analysis and try to explain why we ran the tests that we did
6. Conclusion

# Kruskal Wallis Test-> Opportunity Score
```{r}
# Null Hypothesis: There is no difference in opportunity score between the different time intervals
# Alternative Hypothesis: There is a difference in opportunity score between the different time intervals
kruskal.test(minutes_summary_df$opportunity_score ~ minutes_summary_df$time_only)
```
p-value < 0.05 so we reject the null hypothesis


```{r}
# Null Hypothesis: There is no difference in opportunity score between the different time intervals
# Alternative Hypothesis: There is a difference in opportunity score between the different time intervals
kruskal.test(hourly_summary_df$opportunity_score ~ hourly_summary_df$hour_military)
```

p-value > 0.05 so we cannot reject the null hypothesis. we will no longer analyze the hourly_summary, as we cannot prove there is a difference between such intervals


# Kruskal Wallis Test -> Intensity
```{r}
# Null Hypothesis: There is no difference in intensity between the different time intervals
# Alternative Hypothesis: There is a difference in intensity between the different time intervals
kruskal.test(minutes_summary_df$norm_intensity ~ minutes_summary_df$time_only)
```
p-value < 0.05 so we reject the null hypothesis

# Kruskal Wallis Test -> Sleep
```{r}
# Null Hypothesis: There is no difference in sleep between the different time intervals
# Alternative Hypothesis: There is a difference in sleep between the different time intervals
kruskal.test(minutes_summary_df$norm_sleep ~ minutes_summary_df$time_only)
```
p-value < 0.05 so we reject the null hypothesis


- Opportunity scores, activity intensity, and sleep patterns change significantly at different times, meaning time affects these factors.
- However, when looking at opportunity scores by the hour, there was no big difference. This suggests that analyzing time in smaller units (like minutes) gives better insights. 
- To understand these changes better, we can look at which specific time periods have the biggest impact.


# Chi Square Test Opportunity Score By Hour Intervals
```{r}
# check for chi square assumptions in 10 minute intervals
nrow(minutes_summary_df %>% filter(opportunity_level == "good"))
nrow(minutes_summary_df %>% filter(opportunity_level == "mid"))
nrow(minutes_summary_df %>% filter(opportunity_level == "bad"))
```

All of the expected values are greater than 5, so we can run a chi square test.

# Chi Square Test Opportunity Score By 10 Minutes
```{r}
# Null Hypothesis: There is no association between 10 minute intervals and opportunity level, so the distribution of good, mid, and bad is similar throughout all the 10 minute intervals.
# Alternative Hypothesis: There is an association between 10 minute intervals and opportunity level, so the distribution of good, mid, and bad is not similar throughout all the 10 minute intervals.
chisq.test(minutes_summary_df$time_only, minutes_summary_df$opportunity_level)
```

p < 0.05, so we reject the null hypothesis, and accept the alternative hypothesis. There is an association between 10 minute intervals and opportunity level, so the distribution of good, mid, and bad is not similar throughout all the 10 minute intervals.


- The chi-square test shows that opportunity levels (good, mid, bad) are not evenly distributed across 10-minute intervals. This means that time plays a role in determining opportunity levels. 
- Since the p-value is very small (p < 0.05), we can confidently say that certain time intervals have a higher or lower chance of being classified as good, mid, or bad.
- Further analysis could help identify which specific time periods have the most impact.(Post-Hoc Test)

## Conclusions

In this study, we wanted to figure out the best times to show marketing ads based on when people are most likely to pay attention. We thought that people would be more open to ads when they are awake but resting. To test this, we used Fitbit data from 35 participants over a month, looking at their sleep patterns, activity levels, steps, and calories burned.

Through our analysis, we found that people are more active during the day and less active at night. We created a score that helped us understand the best times to show ads, based on how active or restful the person was. We gave more importance to sleep, since someone who is sleeping won’t be paying attention to ads, while someone who is only lightly active might still see them.

The results showed that early mornings and late evenings were better times for ads, as people were awake but not too active. However, we also found that most of the time slots had lower scores, meaning those periods weren't ideal for ads.

In conclusion, this study suggests that showing ads when people are awake but resting might work better. In the future, we could improve this by considering more factors to make the timing of ads even more accurate.

# Final Slides

What does this mean for the Businesses?

1. Pritorize high scoring time slots to maximize how this analysis can help

To get the most viewers for TV ads, it's important to show them during peak hours when more people are watching. These times are usually in the evening (7–10 PM) when families and working professionals are relaxing at home.
Looking at TV ratings and past viewer data can help find the best time slots to reach the biggest audience

2. Using demographic data to fine tune targeting and improved Ad effectiveness

Different ads work better for different groups of people. Advertisers should look at factors like age, gender, income, and interests to place ads in the right time slots. For example, a toy company should run ads during children's shows, while a luxury brand might target late-night business news programs. Matching ads to the right audience makes them more effective and increases sales.

3. Find the best days of the week to run ads

Audience habits change throughout the week:

-Weekday evenings (Monday-Friday, 7–10 PM): Popular for news and entertainment.

-Weekend mornings and afternoons: Good for family-friendly ads.

-Friday and Saturday nights: Best for ads related to entertainment and sports.

By studying these trends, advertisers can focus their budgets on days when more people are watching and avoid wasting money on low-viewership periods

4. Extend research to seasonal patterns to determine the best time of year for Ad placements

Viewership changes depending on the time of year. Some key seasons include:

-Holiday seasons (Thanksgiving, Christmas, New Year’s): People watch more TV, making it a great time for ads related to shopping, travel, and food.

-Back-to-school season (August-September): A good time for ads about school supplies, clothes, and tech gadgets.

-Big sports events (Super Bowl, World Cup, Olympics): Best for brands targeting sports fans.

-Summer months: People spend more time outdoors, so TV ads may need to focus on families and vacation-related products.

-By adjusting ad placements based on seasonal trends, advertisers can reach the right audience at the right time.