---
title: "writeup.rmd"
author: "Josh Schechter"
date: "2025-03-18"
output: html_document
code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Work on this branch. Do not push your code. We will copy and paste and put the code into one persons local and they will push and merge to develop/
# This is our writeup. Everybody should write at least 1000 words for their section, and also include their plots and the code for that section.

## Import Libraries
```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)
```

## Load Intervals
```{r}
intervals_df <- read.csv("C:/Users/JDS221/GWU/DATS6101/projects/DATS6101_Project/data/fitbit/timestep_data/intervals.csv")
```

## Setup (Josh)

## Time series graph of individual features
```{r}
```

## Correlation Matrix of All Features of Interest
```{r}
```


## Opportunity Score
```{r}
```

## EDA (Lincoln)

## Boxplot
```{r}
```

## Bar Graph
```{r}
```

## QQ Plot
```{r}
```

## Histogram
```{r}
```


## Statistical Analysis (Naiyani)

## Kruskal Wallis
```{r}
```

## Chi Square
# Need to write new code for this
```{r}
```

## Post Hoc
```{r}
```

## Conclusions
# Final Slides




